Ecosystem carbon balance in the Hawaiian Islands under different scenarios of future climate and land use change

### 2021-08-03 Manuscript/journal_correspondence/ERL_responseletter.Rmd

---
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    keep_tex: no
    latex_engine: xelatex
header-includes:
   - \linespread{1.05}
   - \newcommand{\blankline}{\quad\pagebreak[2]}
   - \usepackage{eso-pic}
   - \usepackage{graphicx}
   - \newcommand\BackgroundPic{\put(0,0){\parbox[b][\paperheight]{\paperwidth}{\vfill\centering \includegraphics[width=\paperwidth,height=\paperheight, keepaspectratio]{usgs_letterhead.pdf} \vfill }}}
geometry: top=1in, left=1in, bottom=1in, right=1in
mainfont: Times New Roman
fontsize: 12pt
urlcolor: blue
---

\AddToShipoutPicture*{\BackgroundPic}
\pagenumbering{arabic}

@@

\blankline

Dear Editors:

\blankline

Please consider this revised version of article ERL-111271, "Ecosystem carbon balance in the Hawaiian Islands under different scenarios of climate and land use change", for publication in _Environmental Research Letters_. We have fully addressed each of the referees’ comments in detail below. Our point-by-point replies immediately follow each referee comment, and we include line numbers of the revised manuscript at the end of each response to identify where we made changes to the text.

Long-term model projections such as ours will be critical to understanding how future land use and climate change may interact to influence the achievement of climate mitigation goals, and so we expect our paper to be of immediate interest to a wide audience of carbon cycle scientists, ecosystem modelers, land managers, and policy makers. We hope you find our revisions and responses to referees’ comments sufficient to merit publication of our manuscript.

@@

\blankline

Best regards,

\raggedright
Paul C. Selmants, Ph.D. 
\linebreak
Research Ecologist 
\linebreak
U.S. Geological Survey

## Referee 1, comments to the author:

**The manuscript describes an impressive effort projecting ecosystem carbon balance in the Hawaiian Islands under climate change and land use change scenarios. This assessment of future regional terrestrial carbon dynamics can provide a support to local planning of climate mitigation strategies through ecosystem carbon sequestration. The manuscript is well written and organized. However, more clarification and justification are needed for methodology and assumptions about how future NPP and land cover change is modelled. And also, more details are needed such as maps illustrating how projected change in carbon flux (e.g., NPP and NBP) and carbon pools (live biomass, soil organic soil) is spatially distributed. I have provided detailed comments below.**

**L109, why need to assign each cell a state type? What is consideration for it?**

**L114, transitions between ‘state type’? Is it ‘land cover classes’? According to the definition of state type as unique combination of moisture zones, islands and land cover classes at L110, is there transition from dry forest and mesic forest in the same island?**

**L123-126, high land use scenario samples historical rate between median and maximum, but why not low land use scenario sample rates between minimum and median? What is the justification that low land use scenario will have rate lower than historical records?**

**L126, missing description for urbanization rate for high land use scenario.**

**L133, what transitions can increase forest area such as low land use scenario shown in figure S5? Does tree plantation refer to reforestation or afforestation? If afforestation, what is the source land cover class for this plantation?**

**L136, it will be helpful to add average rotation cycle corresponding to the 75% and 40% historical harvest rate? Like how soon planted trees will be harvested in low and high land use scenarios.**

**L137, will wildfire take place only in forests? If so, what is the resulting land cover class? Would it still recover back to forest after a certain time?**

**L140, is state type land cover classes? Since ‘state type’ is defined as combination of moisture zone, island, and land cover type at line 110, how can ‘state type’ be re-combined with moisture zone and island here?**

**L158, moisture zone is included in state type as defined at line L110, isn’t it?**

**L162, how litter and dead wood pools were initialized?**

 **L161, what is 30-year climate normal come from? Does it have consistent temperature and rainfall distribution with RCP projections during 2010 and 2020? Or is there any bias between 30-year climate normal and RCP projections?**

**L163, figure 2 shows total live biomass in 2010 is about 68 Tg C, but total forest aboveground biomass is estimated as 36 Tg C in Asner et al 2016 which is used for initialization here. Where is the additional 32 Tg C from? What is ratio of aboveground biomass to total live biomass in IBIS?**

**L163, what is calibrated GPP NPP rate? It will be helpful to list out the mean or the range.**

**L166-168, is similar multiplier also applied to living biomass? If not, will this single NPP multiplier break the equilibrium between NPP and biomass? For example, for a simulation cell, NPP and biomass should reach dynamic equilibrium in IBIS simulation, a given NPP could only support carbon turnover of a certain number of trees. External decease NPP by a multiplier will in turn reduce biomass in order to match it, causing artificial reduction in biomass.**

**Line 173, why assume Rh as ratio of NPP? Is the ratio constant over time and across climate change scenarios? Why not to estimate Rh based on soil moisture, temperature and soil carbon stocks? Because of lack of soil temperature in RCP 4.5 and 8.5? Need more justification for this flux ratio method.**

**L177-179, are the growth and decay multipliers varied at annual basis? How are multipliers derived for time periods CMIP5 projections do not cover? Such as period between 2020-2049? It is very necessary to include the distribution of these multipliers, are they enhancing or weakening NPP over time at two RCPs?**

**L182, the empirical model used here to relate NPP to temperature and rainfall is core to projections of carbon balance. It is necessary to present this empirical relationship in the paper, such as what the optimal temperature of NPP is and how it changes with rainfall. For example, figure S4 shows projections of temperature and rain, there should be corresponding projections of NPP change.**

**L184, what is the temporal coverage and resolution of these climate projections? Do them only cover the period of mid-century (2049-2069) and end-of-century (2070-2099) as described in L179?**

**L239: is the NPP decrease during 2010-2020 evident in satellite observations? To what extent of this decrease is caused by climate conditions change rather than by forest area changes?**

**L242. It is misleading to state ‘climate change led to increased Rh over time’ as figure 3 shows decreasing trend of Rh under high land use scenario and two RCP climate change scenarios. It should be revised as something like ‘switching from RCP 4.5 to 8.5 led to increased Rh’.**

**L244, why the increase in total soil carbon stock (figure 2) will not enhance Rh? Is it due to the assumption made by this work that Rh is only ratio of NPP and latter decrease over time? What are drivers of Rh decrease?**

**L244, why Rh shows very sharp decrease before 2020 under all land use scenarios with all RCPs, such decrease rate seems to be higher than the rest period. Are these decreases supported by observations?**

**Line 341, what is the source of ‘Hawai’i-specific climate data’?**

## Referee 2, comments to the author:

**The manuscript entitled “Ecosystem carbon balance in the Hawaiian Islands under different scenarios of future climate and land use change” is interesting and suitable for publication but several mistakes in the manuscript should be corrected before acceptance.**

**Specific comments**

**Authors should carefully recheck the manuscript for the typos and also write the meaning of all acronyms because there are some which are not presented.**

**All the figures are clearly presented but authors must check all the figure pixels again. Some of them are not very clear.**

**I suggest that authors check the bibliographic references in the main text and revise the format of documents to meet the requirement of the Journal.**

**In the abstract, please add more information about the methods of the experiment.**

**In the abstract, please don't write the same conclusion lines as they were written in the main conclusion section.**

**Hypothesis should be given. How this work is different from the available literature?**

**Objectives should be rewritten in detail and comprehensively.**

**How did the authors perform the calibration and validation process?**

**How did they select the sensitive parameters of the model that they used?**

**Discussion, the discussion part is not related to the results part. Why you discuss these? What does it have to do with the result?**

### 2021-08-04 Manuscript/journal_correspondence/ERL_responseletter.Rmd

Dear Editors:

\blankline

Please consider this revised version of article ERL-111271, "Ecosystem carbon balance in the Hawaiian Islands under different scenarios of climate and land use change", for publication in _Environmental Research Letters_. We have fully addressed each of the referees’ comments in detail below. Our point-by-point replies immediately follow each referee comment, and we include line numbers of the revised manuscript at the end of each response to identify where we made changes to the text.

Long-term model projections such as ours will be critical to understanding how future land use and climate change may interact to influence the achievement of climate mitigation goals, and so we expect our paper to be of immediate interest to a wide audience of carbon cycle scientists, ecosystem modelers, land managers, and policy makers. We hope you find the revisions and responses to referees’ comments sufficient to merit publication of our manuscript.

@@

\raggedright
Paul C. Selmants, Ph.D. 
\linebreak
Research Ecologist 
\linebreak
U.S. Geological Survey

## Referee 1, comments to the author:

**The manuscript describes an impressive effort projecting ecosystem carbon balance in the Hawaiian Islands under climate change and land use change scenarios. This assessment of future regional terrestrial carbon dynamics can provide a support to local planning of climate mitigation strategies through ecosystem carbon sequestration. The manuscript is well written and organized. However, more clarification and justification are needed for methodology and assumptions about how future NPP and land cover change is modelled. And also, more details are needed such as maps illustrating how projected change in carbon flux (e.g., NPP and NBP) and carbon pools (live biomass, soil organic soil) is spatially distributed. I have provided detailed comments below.**

**L109, why need to assign each cell a state type? What is consideration for it?**  
The state of each cell tracks a range of information about that cell, including land cover class, moisture zone, island, age, and time since transition. This set of information defines the probabilities that each cell can undergo a particular type of transition, as well as how often each of these transitions can occur. We have added text and references to this section and the next that better describes the state of each cell (lines ).

**L114, transitions between ‘state type’? Is it ‘land cover classes’? According to the definition of state type as unique combination of moisture zones, islands and land cover classes at L110, is there transition from dry forest and mesic forest in the same island?**  
State type in this model is the unique combination of moisture zone, island, and land cover class, as stated on line 110 of the original manuscript. This information affects transition probabilities, such that dry grassland on Hawaiʻi Island has a different probability of transitioning to agriculture than mesic grassland on Oʻahu. We have added text that better defines how state type influences transition probabilities ().

**L123-126, high land use scenario samples historical rate between median and maximum, but why not low land use scenario sample rates between minimum and median? What is the justification that low land use scenario will have rate lower than historical records?**  
To quantify the impact of a low land use scenario, rates of land use change were sampled from the lower range of historical rates (median and below, bounded by zero). We have added text that justifies land use change rates sampled from a range lower than historical rates in the low land use scenario (lines xxx - xxx).

**L126, missing description for urbanization rate for high land use scenario.**  
As the referee themselves point this out in their previous comment, transition rates in the high land use scenario were sampled from uniform distributions bounded by the median and maximum historical rates of agricultural expansion, agricultural contraction, and urbanization for each island. This description is on lines 121-123 of the original manuscript and lines xxx - xxxx of the revised manuscript. No changes were made in response to this comment.

@@

**L133, what transitions can increase forest area such as low land use scenario shown in figure S5? Does tree plantation refer to reforestation or afforestation? If afforestation, what is the source land cover class for this plantation?**

**L136, it will be helpful to add average rotation cycle corresponding to the 75% and 40% historical harvest rate? Like how soon planted trees will be harvested in low and high land use scenarios.**  
We have added text defining the historical harvest rate as (lines xxx - xxxx).

**L137, will wildfire take place only in forests? If so, what is the resulting land cover class? Would it still recover back to forest after a certain time?**  
Wildland fire in Hawaiʻi takes place primarily in non-native grasslands, but can also occur in shrublands and forest. We have added text that describes where wildland fire can occur in the Hawaiian Islands (lines xxx - xxx).

**L140, is state type land cover classes? Since ‘state type’ is defined as combination of moisture zone, island, and land cover type at line 110, how can ‘state type’ be re-combined with moisture zone and island here?**  
We agree that this was a confusing use of terminology. We have changed "state type" to "land cover class" (line xxx).

**L158, moisture zone is included in state type as defined at line L110, isn’t it?**  
We agree this was, again, a confusing use of terminology. We have omitted "moisture zone" and now only refer to "state type" in this sentence (line xxx).

@@

**L162, how litter and dead wood pools were initialized?**

 **L161, what is 30-year climate normal come from? Does it have consistent temperature and rainfall distribution with RCP projections during 2010 and 2020? Or is there any bias between 30-year climate normal and RCP projections?**  
 The 30-year climate normals for precipitation and temperature come from the references cited (Giambelluca et al. 2013, 2014).

**L163, figure 2 shows total live biomass in 2010 is about 68 Tg C, but total forest aboveground biomass is estimated as 36 Tg C in Asner et al 2016 which is used for initialization here. Where is the additional 32 Tg C from? What is ratio of aboveground biomass to total live biomass in IBIS?**  
Asner et al. (2016) only quantifies aboveground forest biomass. The additional 32 Tg C comes from belowground forest biomass, as well as total live biomass in grasslands, shrublands, and agriculture. We have added text describing how we estimated belowground forest biomass.

**L163, what is calibrated GPP NPP rate? It will be helpful to list out the mean or the range.**
We have added text listing the range and median rate of GPP from the calibration dataset (Kimball et al. 2017) and added a table of mean NPP rates from IBIS for each moisture zone and land cover class (xxx).

**L166-168, is similar multiplier also applied to living biomass? If not, will this single NPP multiplier break the equilibrium between NPP and biomass? For example, for a simulation cell, NPP and biomass should reach dynamic equilibrium in IBIS simulation, a given NPP could only support carbon turnover of a certain number of trees. External decease NPP by a multiplier will in turn reduce biomass in order to match it, causing artificial reduction in biomass.**  
Biomass already accumulated will not be reduced if NPP declines. Rather, the rate of biomass increase will decline as NPP declines. This is true in our model, as it is in the real world. We have added text that better describes how and why we apply the NPP spatial multiplier (xxx).

@@

**Line 173, why assume Rh as ratio of NPP? Is the ratio constant over time and across climate change scenarios? Why not to estimate Rh based on soil moisture, temperature and soil carbon stocks? Because of lack of soil temperature in RCP 4.5 and 8.5? Need more justification for this flux ratio method.**

**L177-179, are the growth and decay multipliers varied at annual basis? How are multipliers derived for time periods CMIP5 projections do not cover? Such as period between 2020-2049? It is very necessary to include the distribution of these multipliers, are they enhancing or weakening NPP over time at two RCPs?**

**L182, the empirical model used here to relate NPP to temperature and rainfall is core to projections of carbon balance. It is necessary to present this empirical relationship in the paper, such as what the optimal temperature of NPP is and how it changes with rainfall. For example, figure S4 shows projections of temperature and rain, there should be corresponding projections of NPP change.**  
Projections of NPP change by scenario are presented in Figure 3.

**L184, what is the temporal coverage and resolution of these climate projections? Do them only cover the period of mid-century (2049-2069) and end-of-century (2070-2099) as described in L179?**  
Yes, climate projections only cover the period of mid-century (2049) and end-of-century (2070-2099) as described on line 179 of the original manuscript. We have added text that described how we

### 2021-08-06 Manuscript/HI_Model_ms.Rmd
updated manuscript file, supplementary file, and response letter file.

## Abstract
The State of Hawaiʻi passed legislation to be carbon neutral by 2045, a goal that will partly depend on carbon sequestration by terrestrial ecosystems. However, there is considerable uncertainty surrounding the future direction and magnitude of the land carbon sink in the Hawaiian Islands. We used simulation modeling to assess how projected future changes in climate and land use will influence ecosystem carbon balance in the Hawaiian Islands under all combinations of two radiative forcing scenarios (RCP 4.5 and RCP 8.5) and two land use scenarios (low and high) over a 90-year timespan from 2010-2100. Collectively, terrestrial ecosystems of the Hawaiian Islands acted as a net carbon sink under low radiative forcing (RCP 4.5) for the entire 90-year simulation period, with low land use change further enhancing carbon sink strength. In contrast, Hawaiian terrestrial ecosystems transitioned from a net sink to a net source of CO~2~ to the atmosphere under high radiative forcing (RCP 8.5), with high land use accelerating this transition and exacerbating net carbon loss. A sensitivity test of the CO~2~ fertilization effect on plant productivity revealed it to be a major source of uncertainty in projections of ecosystem carbon balance. Greater mechanistic understanding of plant productivity responses to rising atmospheric CO~2~ will be essential to realistically constrain simulation models used to evaluate the effectiveness of ecosystem-based climate mitigation strategies.

@@

## 2. Methods
We used the Land Use and Carbon Scenario Simulator (LUCAS), an integrated landscape change and carbon gain-loss model, to project changes in ecosystem carbon balance for the seven main Hawaiian Islands. The landscape change portion of LUCAS is a state-and-transition model, where 'state' is the pre-defined information specific to each simulation cell (e.g., land cover class, climate zone) and 'transition' is the change in state of each simulation cell over time [@daniel_state-and-transition_2016]. Landscape change occurs by applying a Monte Carlo approach to track the state type and age of each simulation cell in response to a pre-determined set of transition pathways [@daniel_state-and-transition_2016]. The carbon gain-loss portion tracks carbon stocks within each simulation cell over time as continuous state variables, along with a pre-defined set of continuous flows specifying rates of change in stock levels over time [@daniel_integrating_2018; @sleeter_effects_2019]. We parameterized the Hawaiʻi LUCAS model to estimate annual changes in carbon stocks and fluxes in response to land use, land use change, wildland fire, and long-term climate variability for the time period 2010-2100.

@@

### *2.2 States and transitions*
We developed two land use scenarios (low and high) with transition pathways modified from @daniel_state-and-transition_2016. Transition pathways define which state types can be converted to any other state type. Transition pathways between state types were pre-defined to represent urbanization, agricultural contraction, agricultural expansion, harvesting of tree plantations, and wildfire. Agriculture, forest, grassland, shrubland, and tree plantation state types each had multiple transition pathways, i.e., they could each transition to multiple other state types. The barren state type could only transition to developed, so its sole transition pathway was urbanization. Although most state types had an urbanization transition pathway, there was no transition pathway out of an urbanized (developed) state. Water and wetland state types remained static throughout the simulation period.

 Transition targets set the area to be transitioned over time. These targets were based on historical trends of land use change in the Hawaiian Islands from 1992-2011 [@noaa_coastal_2020] and on population projections for the State of Hawaiʻi [@kim_population_2018]. For the high land use scenario, transition rates for each timestep and Monte Carlo realization were sampled from uniform distributions bounded by the median and maximum historical rates of agricultural contraction, agricultural expansion, and urbanization for each island. For the low land use scenario, rates of agricultural contraction and expansion were sampled from uniform distributions bounded by zero and the minimum historical rates for each island. Urbanization rates in the low land use scenario were based on island-level population estimates and projections at five year intervals from 2010-2045 [@kim_population_2018]. We converted population projections into urbanization transition targets following @sleeter_future_2017 by calculating population density for each island and then projecting future developed area based on the five-year incremental change in island population. The spatial extent of agricultural contraction, agricultural expansion, and urbanization was constrained in both land use scenarios based on existing zoning maps [@daniel_state-and-transition_2016]. Transition targets for tree plantation harvest were set at ~75% of recent historical rates in the high land use scenario and ~40% of recent historical rates in the low land use scenario [@daniel_state-and-transition_2016]. In both land use scenarios, approximately 60% of tree plantation harvests were replacement harvests resulting in conversion to agriculture. The remaining 40% were rotation harvests replanted to *Eucalyptus* spp., which represents the vast majority of plantation forestry in the State of Hawaiʻi.

The wildfire transition sub-model was modified from @daniel_state-and-transition_2016 by incorporating a new 21-year historical wildfire spatial database of the Hawaiian Islands (supplemental figure 2). We used this new spatial database to calculate historical wildfire size distribution and ignition probabilities for each unique combination of moisture zone (supplemental figure 1), island, and state type (figure 1) for the years 1999-2019. Starting in 2020, the number and size of fires was randomly drawn from one of these historical year-sets for each timestep and Monte Carlo realization, using burn severity probabilities from @selmants_baseline_2017. Wildfire in the low land use scenario was sampled from the subset of historical fire years at or below the median area burned statewide from 1999-2019. The high land use scenario sampled from historical fire years above the median area burned over the same 21-year period (supplemental figure 2a).

@@

 Initial carbon stocks and baseline carbon flows were estimated based on the moisture zone (supplemental figure 1), state type, and age of each simulation cell using a lookup table derived from the Integrated Biosphere Simulator [IBIS; @foley_integrated_1996; @liu_critical_2020], a process-based dynamic global vegetation model. We initiated IBIS with minimal vegetation and simulated forward for 110 years using 30-year climate normals for the Hawaiian Islands [@giambelluca_online_2013; @giambelluca_evapotranspiration_2014]. We calibrated IBIS carbon stocks with statewide gridded datasets of soil organic carbon [@ssurgo_2016] and forest aboveground live biomass [@asner_rapid_2016]. We also calibrated gross photosynthesis in IBIS using a Hawaiʻi-specific gridded dataset derived from MODIS satellite imagery [@kimball_evaluating_2017].

 Net primary production for each simulation cell was calculated as the mean IBIS-derived value for each combination of moisture zone and state type adjusted with a spatially explicit stationary growth multiplier to reflect local variation driven by microclimate [@sleeter_effects_2019]. We calculated this spatial growth multiplier as the NPP anomaly for each simulation cell relative to mean NPP values for each combination of moisture zone and state type based on empirical relationships between total annual NPP and mean annual rainfall or temperature [@schuur_productivity_2003; @del_grosso_global_2008] using Hawaiʻi-specific climate data [@giambelluca_online_2013; @giambelluca_evapotranspiration_2014]. Soil carbon flux to the atmosphere (R~h~) and aquatic soil carbon losses (leaching and overland flow) were estimated as the ratio of the IBIS-derived flux for each combination of moisture zone and state type to the microclimate-adjusted NPP value for each simulation cell. All other carbon flow rates were estimated as the ratio of the mean IBIS-derived flux for each combination of moisture zone and state type to the size of the originating carbon stock at each age [@sleeter_effects_2018; @daniel_integrating_2018].

>>>>

## Abstract
The State of Hawaiʻi passed legislation to be carbon neutral by 2045, a goal that will partly depend on carbon sequestration by terrestrial ecosystems. However, there is considerable uncertainty surrounding the future direction and magnitude of the land carbon sink in the Hawaiian Islands. We used the Land Use and Carbon Scenario Simulator (LUCAS), a spatially explicit stochastic  simulation model that integrates landscape change and carbon gain-loss, to assess how projected future changes in climate and land use will influence ecosystem carbon balance in the Hawaiian Islands under all combinations of two radiative forcing scenarios (RCPs 4.5 and 8.5) and two land use scenarios (low and high) over a 90-year timespan from 2010-2100. We used simulation modeling to assess how projected future changes in climate and land use will influence ecosystem carbon balance in the Hawaiian Islands under all combinations of two radiative forcing scenarios (RCP 4.5 and RCP 8.5) and two land use scenarios (low and high) over a 90-year timespan from 2010-2100. Collectively, terrestrial ecosystems of the Hawaiian Islands acted as a net carbon sink under low radiative forcing (RCP 4.5) for the entire 90-year simulation period, with low land use change further enhancing carbon sink strength. In contrast, Hawaiian terrestrial ecosystems transitioned from a net sink to a net source of CO~2~ to the atmosphere under high radiative forcing (RCP 8.5), with high land use accelerating this transition and exacerbating net carbon loss. A sensitivity test of the CO~2~ fertilization effect on plant productivity revealed it to be a major source of uncertainty in projections of ecosystem carbon balance. Greater mechanistic understanding of plant productivity responses to rising atmospheric CO~2~ will be essential to realistically constrain simulation models used to evaluate the effectiveness of ecosystem-based climate mitigation strategies.

@@

## 2. Methods
We used the Land Use and Carbon Scenario Simulator (LUCAS), an integrated landscape change and carbon gain-loss model, to project changes in ecosystem carbon balance for the seven main Hawaiian Islands. The landscape change portion of LUCAS is a state-and-transition model, where 'state' is the pre-defined information specific to each simulation cell (e.g., land cover class, climate zone) and 'transition' is the change in state of each simulation cell over time [@daniel_state-and-transition_2016]. Landscape change occurs by applying a Monte Carlo approach to track the state type and age of each simulation cell in response to a pre-determined set of transition pathways [@daniel_state-and-transition_2016]. Transition pathways define which state types can be converted to any other state type. The carbon gain-loss portion tracks carbon stocks within each simulation cell over time as continuous state variables, along with a pre-defined set of continuous flows specifying rates of change in stock levels over time [@daniel_integrating_2018; @sleeter_effects_2019]. We parameterized the Hawaiʻi LUCAS model to estimate annual changes in carbon stocks and fluxes in response to land use, land use change, wildland fire, and long-term climate variability for the time period 2010-2100.

@@

### *2.2 States and transitions*
We developed two land use scenarios (low and high) with transition pathways modified from @daniel_state-and-transition_2016. Transition pathways were pre-defined to represent urbanization, agricultural contraction, agricultural expansion, harvesting of tree plantations, and wildfire. Agriculture, forest, grassland, shrubland, and tree plantation land cover classes each had multiple transition pathways, i.e., they could each be converted into multiple other land cover classes. The barren land cover class could only transition to developed, so its sole transition pathway was urbanization. Although most land cover classes had an urbanization transition pathway, there was no transition pathway out of an urbanized (developed) state. Water and wetland land cover classes remained static throughout the simulation period.

 Transition targets set the area to be transitioned over time. These targets were based on historical trends of land use change in the Hawaiian Islands from 1992-2011 [@noaa_coastal_2020] and on population projections for the State of Hawaiʻi [@kim_population_2018]. For the high land use scenario, transition rates for each timestep and Monte Carlo realization were sampled from uniform distributions bounded by the median and maximum historical rates of agricultural contraction, agricultural expansion, and urbanization for each island. For the low land use scenario, rates of agricultural contraction and expansion were sampled from uniform distributions bounded by zero and the minimum historical rates for each island. Urbanization rates in the low land use scenario were based on island-level population estimates and projections at five year intervals from 2010-2045 [@kim_population_2018]. We converted population projections into urbanization transition targets following @sleeter_future_2017 by calculating population density for each island and then projecting future developed area based on the five-year incremental change in island population. The spatial extent of agricultural contraction, agricultural expansion, and urbanization was constrained in both land use scenarios based on existing zoning maps [@daniel_state-and-transition_2016]. Tree plantation forestry in the State of Hawaiʻi is primarily short-rotation (5-7 years) *Eucalyptus* spp., harvested at a rate of approximately 2 km^2^ y^-1^ over the past decade. Transition targets for tree plantation harvest were set at ~75% of recent historical rates in the high land use scenario and ~40% of recent historical rates in the low land use scenario [@daniel_state-and-transition_2016]. In both land use scenarios, approximately 60% of tree plantation harvests were replacement harvests resulting in conversion to agriculture. The remaining 40% were rotation harvests replanted to *Eucalyptus* spp. No land cover class could be converted to tree plantation in either land use scenario, consistent with recent historical trends of stable or declining area of tree plantations.

The wildfire transition sub-model was modified from @daniel_state-and-transition_2016 by incorporating a new 21-year historical wildfire spatial database of the Hawaiian Islands (supplemental figure 2). We used this new spatial database to calculate historical wildfire size distribution and ignition probabilities for each unique combination of moisture zone (supplemental figure 1), island, and land cover class (figure 1) for the years 1999-2019. Starting in 2020, the number and size of fires was randomly drawn from one of these historical year-sets for each timestep and Monte Carlo realization, using burn severity probabilities from @selmants_baseline_2017. Wildfire in the low land use scenario was sampled from the subset of historical fire years at or below the median area burned statewide from 1999-2019. The high land use scenario sampled from historical fire years above the median area burned over the same 21-year period (supplemental figure 2a). The vast majority of wildland fire in Hawaiʻi is human-caused and initiates in non-native grasslands and shrublands [@trauernicht_contemporary_2015], but can spread into nearby forest areas. Moderate to high severity fires that spread into mesic and wet forests can convert these areas into grasslands about half the time. However, moderate to high severity fires in mesic and wet forests account for <5% of the total annual area burned, on average [@selmants_baseline_2017].

@@

 Initial carbon stocks and baseline carbon flows were estimated based on the state type and age of each simulation cell using a lookup table derived from the Integrated Biosphere Simulator [IBIS; @foley_integrated_1996; @liu_critical_2020], a process-based dynamic global vegetation model. We initiated IBIS with minimal vegetation and simulated forward for 110 years using 30-year climate normals for the Hawaiian Islands [@giambelluca_online_2013; @giambelluca_evapotranspiration_2014]. We calibrated IBIS carbon stocks with statewide gridded datasets of soil organic carbon [@ssurgo_2016] and forest aboveground live biomass [@asner_rapid_2016]. We also calibrated gross photosynthesis in IBIS using a Hawaiʻi-specific gridded dataset derived from MODIS satellite imagery [@kimball_evaluating_2017].

 Net primary production for each simulation cell was calculated as the mean IBIS-derived value for each combination of moisture zone and land cover class (supplemental Table 1) adjusted with a spatially explicit stationary growth multiplier to reflect local variation driven by microclimate [supplemental Figure 4; @sleeter_effects_2019]. We calculated this spatial growth multiplier as the NPP anomaly for each simulation cell relative to mean NPP values for each combination of moisture zone and land cover class based on empirical relationships between total annual NPP and mean annual rainfall or temperature [@schuur_productivity_2003; @del_grosso_global_2008] using Hawaiʻi-specific climate data [@giambelluca_online_2013; @giambelluca_evapotranspiration_2014]. Soil carbon flux to the atmosphere (R~h~) and aquatic soil carbon losses (leaching and overland flow) were estimated as the ratio of the IBIS-derived flux for each combination of moisture zone and state type to the microclimate-adjusted NPP value for each simulation cell. All other carbon flow rates were estimated as the ratio of the mean IBIS-derived flux for each combination of moisture zone and state type to the size of the originating carbon stock at each age [@sleeter_effects_2018; @daniel_integrating_2018].

### 2021-08-06 Manuscript/HI_Model_ms20210405.Rmd
updated manuscript file, supplementary file, and response letter file.

 Transition targets were based on historical trends of land use change in the Hawaiian Islands from 1992-2011 [@noaa_coastal_2020] and on population projections for the State of Hawaiʻi [@kim_population_2018]. For the high land use scenario, transition rates for each timestep and Monte Carlo realization were sampled from uniform distributions bounded by the median and maximum historical rates of agricultural contraction, agricultural expansion, and urbanization for each island. For the low land use scenario, rates of agricultural contraction and expansion were sampled from uniform distributions bounded by zero and the minimum historical rates for each island. Urbanization rates in the low land use scenario were based on island-level population estimates and projections at five year intervals from 2010-2045 [@kim_population_2018]. We converted population projections into urbanization transition targets following @sleeter_future_2017 by calculating population density for each island and then projecting future developed area based on the five-year incremental change in island population. The spatial extent of agricultural contraction, agricultural expansion, and urbanization was constrained in both land use scenarios based on existing zoning maps [@daniel_state-and-transition_2016]. Transition targets for tree plantation harvest were set at ~75% of recent historical rates in the high land use scenario and ~40% of recent historical rates in the low land use scenario [@daniel_state-and-transition_2016]. In both land use scenarios, approximately 60% of tree plantation harvests were replacement harvests resulting in conversion to agriculture. The remaining 40% were rotation harvests replanted to *Eucalyptus* spp.

>>>>

 Transition targets were based on historical trends of land use change in the Hawaiian Islands from 1992-2011 [@noaa_coastal_2020] and on population projections for the State of Hawaiʻi [@kim_population_2018]. For the high land use scenario, transition rates for each timestep and Monte Carlo realization were sampled from uniform distributions bounded by the median and maximum historical rates of agricultural contraction, agricultural expansion, and urbanization for each island. For the low land use scenario, rates of agricultural contraction and expansion were sampled from uniform distributions bounded by zero and the minimum historical rates for each island. Setting the lower boundary of agricultural contraction and expansion rates at zero in the low land use scenario allows for the possibility that agricultural land area will stabilize over time, because one or more timesteps within a given Monte Carlo iteration may experience no change in agricultural land area. We consider this to be a reasonable possibility under a low land use change scenario, but the vast majority of timesteps and Monte Carlo iterations will experience changes in agricultural land area within the historical range. Urbanization rates in the low land use scenario were based on island-level population estimates and projections at five year intervals from 2010-2045 [@kim_population_2018]. We converted population projections into urbanization transition targets following @sleeter_future_2017 by calculating population density for each island and then projecting future developed area based on the five-year incremental change in island population. The spatial extent of agricultural contraction, agricultural expansion, and urbanization was constrained in both land use scenarios based on existing zoning maps [@daniel_state-and-transition_2016]. Transition targets for tree plantation harvest were set at ~75% of recent historical rates in the high land use scenario and ~40% of recent historical rates in the low land use scenario [@daniel_state-and-transition_2016]. In both land use scenarios, approximately 60% of tree plantation harvests were replacement harvests resulting in conversion to agriculture. The remaining 40% were rotation harvests replanted to *Eucalyptus* spp.
